Limitations on the Two-Infall Scenario in the Era of Large Stellar Age Catalogs

### 2025-03-28 src/tex/ms.tex

In detail, we calculate the star formation efficiency (SFE) timescale $\tau_\star\equiv\Sigma_g/\dot\Sigma_\star$ according to the following:
\begin{equation}
    \label{eq:sf-law}
    \tau_\star = 
    \begin{cases}
        \varepsilon(t) \tau_{\rm mol}(t),   & \Sigma_g \ge \Sigma_{g,0} \\
        \varepsilon(t) \tau_{\rm mol}(t) \Big(\frac{\Sigma_g}{\Sigma_{g,0}}\Big)^{-1/2}, & \Sigma_g < \Sigma_{g,0}
    \end{cases}
\end{equation}
where $\Sigma_{g,0} = 10^8\,{\rm M}_\odot\,{\rm kpc}^{-2}$ and $\tau_{\rm mol}(t)=\tau_{\rm mol,0}(t/t_0)^\gamma$, with $\gamma=1/2$, $t_0=13.8\,{\rm Gyr}$ \todo{(?)} and $\tau_{\rm mol,0}=2\,{\rm Gyr}$ \citet{leroy_star_2008}. Previous two-infall studies \citep[e.g.,][]{nissen_high-precision_2020} have adopted a higher SFE during the first infall epoch than during the second, which we emulate through the pre-factor $\varepsilon$:
\begin{equation}
    \label{eq:sfe-prefactor}
    \varepsilon(t) = 
    \begin{cases}
        0.5, & t < t_{\rm max} \\
        1.0, & t \ge t_{\rm max}.
    \end{cases}
\end{equation}
A lower value of $\varepsilon(t<t_{\rm max})$ leads to more efficient star formation during the first infall epoch. Figure \ref{fig:sfe-prefactor} illustrates that this pre-factor largely affects the metallicity of the high-$\alpha$ sequence, with a smaller $\varepsilon$ producing faster enrichment during the first infall and stronger dilution after $t_{\rm max}$. The pre-factor has virtually no effect on the overall [O/Fe] distribution because the model is normalized to produce the same thick-to-thin-disk mass ratio regardless of the details of the star formation law. We adopt $\varepsilon(t<t_{\rm max})=0.5$ for consistency with the two-infall literature. \todo{I actually think I can ditch the pre-factor, since the molecular timescale already decreases with lookback time. In fact, $\tau_{\rm mol}(4.2\,{\rm Gyr}) / \tau_{\rm mol}(13.2\,{\rm Gyr}) \approx 0.5$, which is pretty much the correct ratio to match the two-infall literature. The only catch is that $\tau_{\rm mol}$ is a very smooth transition, while with the prefactor it's an abrupt transition.}

@@

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{src/tex/figures/ofe_distribution_yields.pdf}
    \caption{Stellar [O/Fe] distributions produced by multi-zone models which assume different yield sets and outflow mass-loading factors. Each row presents stellar distributions within a range of absolute midplane distance $|z|$ reported on the far right. Within each panel, the distributions are color-coded according to the bin in galactocentric radius $R_{\rm gal}$ from which they are drawn. Each distribution is normalized so that the area under the curve is 1, and the vertical scale is consistent across each row. The median APOGEE abundance uncertainties are forward-modeled onto the model outputs (see Table \ref{tab:uncertainties}). For visual clarity, each distribution is smoothed with a box-car of width 0.05 dex. {\bf Key takeaway:} The two-infall model produces an intermediate-[O/Fe] peak that is especially prominent in the \yZ{2} model at mid to high latitudes.}
    \label{fig:ofe-distribution-yields}
\end{figure}

>>>>

In detail, we calculate the star formation efficiency (SFE) timescale $\tau_\star\equiv\Sigma_g/\dot\Sigma_\star$ according to the following:
\begin{equation}
    \label{eq:sf-law}
    \tau_\star = 
    \begin{cases}
        \varepsilon(t) \tau_{\rm mol}(t),   & \Sigma_g \ge \Sigma_{g,0} \\
        \varepsilon(t) \tau_{\rm mol}(t) \Big(\frac{\Sigma_g}{\Sigma_{g,0}}\Big)^{-1/2}, & \Sigma_g < \Sigma_{g,0}
    \end{cases}
\end{equation}
where $\Sigma_{g,0} = 10^8\,{\rm M}_\odot\,{\rm kpc}^{-2}$ and $\tau_{\rm mol}(t)=\tau_{\rm mol,0}(t/t_0)^\gamma$, with $\gamma=1/2$, $t_0=13.8\,{\rm Gyr}$ \todo{(?)} and $\tau_{\rm mol,0}=2\,{\rm Gyr}$ \citet{leroy_star_2008}. Previous two-infall studies \citep[e.g.,][]{nissen_high-precision_2020} have adopted a higher SFE during the first infall epoch than during the second, which we emulate through the pre-factor $\varepsilon$:
\begin{equation}
    \label{eq:sfe-prefactor}
    \varepsilon(t) = 
    \begin{cases}
        0.5, & t < t_{\rm max} \\
        1.0, & t \ge t_{\rm max}.
    \end{cases}
\end{equation}
A lower value of $\varepsilon(t<t_{\rm max})$ leads to more efficient star formation during the first infall epoch. Figure \ref{fig:sfe-prefactor} illustrates that this pre-factor largely affects the metallicity of the high-$\alpha$ sequence, with a smaller $\varepsilon$ producing faster enrichment during the first infall and stronger dilution after $t_{\rm max}$. The pre-factor has virtually no effect on the overall [O/Fe] distribution because the model is normalized to produce the same thick-to-thin-disk mass ratio regardless of the details of the star formation law, but a lower value of $\varepsilon$ does narrow the MDF by $\sim0.1$ dex in [Fe/H]. We adopt $\varepsilon(t<t_{\rm max})=0.5$ for consistency with the two-infall literature. We have tested eliminating either $\varepsilon(t)$ or $\tau_{\rm mol}(t)$ from our SFE prescription in multi-zone models and found no substantial difference to our results.
% \todo{I actually think I can ditch the pre-factor, since the molecular timescale already decreases with lookback time. In fact, $\tau_{\rm mol}(4.2\,{\rm Gyr}) / \tau_{\rm mol}(13.2\,{\rm Gyr}) \approx 0.5$, which is pretty much the correct ratio to match the two-infall literature. The only catch is that $\tau_{\rm mol}$ is a very smooth transition, while with the prefactor it's an abrupt transition.}

@@

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{src/tex/figures/ofe_distribution_yields.pdf}
    \caption{Normalized stellar [O/Fe] distributions produced by multi-zone models which assume different yield sets and outflow mass-loading factors. Each row presents stellar distributions within a range of absolute midplane distance $|z|$ reported on the far right, and the vertical scale is consistent across each row. Within each panel, the distributions are color-coded according to the bin in galactocentric radius $R_{\rm gal}$ from which they are drawn. The median APOGEE abundance uncertainties are forward-modeled onto the model outputs (see Table \ref{tab:uncertainties}). For visual clarity, each distribution is smoothed with a box-car of width 0.05 dex. {\bf Key takeaway:} The two-infall model produces an intermediate-[O/Fe] peak that is especially prominent in the \yZ{2} model at mid to high latitudes.}
    \label{fig:ofe-distribution-yields}
\end{figure}
